---
title: "Titanic exercise"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
setwd('C:\\Users\\Chegwood\\Documents\\Data Science Stuff\\Springboard - Foundations of Data Science\\Titanic ggplot exercise')
titanic <- read_csv('titanic_clean_2.csv')
```

## Titanic Exercise from ggplot Tutorial

This is the code that I wrote for the gpplot Tutorial from DataCamp. I used the clean data from the Titanic data wrangling exercies (w/ average ages by class, sex, and survival) so some additional code was needed to produce the graphs from the tutorial.

First, I looked at a bar chart to look at the passengers on board, by class and sex:

```{r}
ggplot(na.omit(titanic[ , c('pclass', 'sex', 'survived', 'age')]), aes(x = factor(pclass), fill = factor(sex))) +
  geom_bar(position = 'dodge')
```

### Faceting by Survival

I then looked to see whether there was a difference if the data was faceted by 'survival':

```{r}
ggplot(na.omit(titanic[ , c('pclass', 'sex', 'survived', 'age')]), aes(x = factor(pclass), fill = factor(sex))) +
  geom_bar(position = 'dodge') +
  facet_grid(. ~ survived)  
```

### Adding in Age

It looks like females were more likely to survive, as well as those in 1st class. Does age have any difference?

First, I set a variable to define a standard 'jitter' width:

```{r, echo = FALSE}
posn_j <- position_jitter(width = 0.5)
```

And finally created a jitter plot with Age as the y axis:

```{r}
ggplot(na.omit(titanic[ , c('pclass', 'sex', 'survived', 'age')]), aes(x = factor(pclass), y = age, col = factor(sex))) +
  geom_jitter(size = 3, alpha = 0.5, position = posn_j) +
  facet_grid(. ~ survived)
```

## Let's Dig a little Deeper

###Floating helps

Here we can see that the vast majority of those who got on a boat survived, while very few who couldn't make it on a boat lived:

```{r, include = FALSE}
titanic_2 <- titanic %>%
  mutate(boat_binary = ifelse(boat == 'None', 0, 1))
```

```{r}
ggplot(na.omit(titanic_2[ , c('pclass', 'sex', 'survived', 'age', 'boat_binary')]), 
       aes(x = age, y = factor(boat_binary), col = factor(pclass), shape = factor(sex))) +
  geom_jitter(size = 3, alpha = 0.5, position = posn_j) +
  facet_grid(survived ~ .)
```

###Proportions vs. Raw Counts
```{r}
ggplot(na.omit(titanic[ , c('pclass', 'sex', 'survived', 'age')]), aes(x = factor(pclass), fill = factor(survived))) +
  geom_bar(position = 'fill') +
  facet_grid(. ~ sex) +
  labs(x = 'Class', y = 'Proportion', fill = 'Survived?')
```

###Relative Tiers from Exit (by deck)

This is a little research and manual effort on my part, but I wanted to see whether being in a cabin close to a staircase and a deck close to the top was helpful to survival. Using blueprints from [link](https://www.encyclopedia-titanica.org/titanic-deckplans/) I created manual tiers of 'closeness' to the top deck. Each deck is a radiates downwards (with A being 1, B being 2... G being 7). All passangers without a cabin listed or with a cabin of T (which I mapped to the Tank deck) was given 8.

The class and deck obviously align quite closely; especially the fact that Class 3 passengers seem entirely relegated to Deck Number 8 (or no cabin at all). Those with cabins, especially those with first class cabins seemed more likely to survive than the others.

```{r include = FALSE}
titanic_2 <- titanic_2 %>% 
  separate(cabin, into = c('deck', 'cabin'), sep = 1) %>% 
  mutate(deck_number = ifelse(deck == 'A', 1, 
          ifelse(deck == 'B', 2, 
          ifelse(deck == 'C', 3, 
          ifelse(deck == 'D', 4,
          ifelse(deck == 'E', 5,
          ifelse(deck == 'F', 6,
          ifelse(deck == 'G', 7, 8))))))))
```

```{r}
ggplot(na.omit(titanic_2[ , c('pclass', 'sex', 'survived', 'age', 'deck_number')]), aes(x = factor(deck_number), y = age, 
                                                                                      col = factor(pclass))) +
  geom_jitter(position = posn_j, alpha = 0.5) +
  facet_grid(. ~ survived) +
  labs(x = 'Deck Number', y = 'Age', col = 'Class')
```

###Survival Rates by Gender, by Class, and by Age Range (in 10 year increments)

```{r include = FALSE}
titanic_2 <- titanic_2 %>% 
  mutate(age_range = cut(titanic_2$age,seq(0, 100, by = 10)))
```

```{r}
titanic_2 %>% 
  group_by(sex) %>% 
  summarise(mean(survived, na.rm = TRUE))

titanic_2 %>% 
  group_by(pclass) %>% 
  summarise(mean(survived, na.rm = TRUE))

titanic_2 %>% 
  group_by(age_range) %>% 
  summarise(mean(survived, na.rm = TRUE))
```

This high-level data analysis shows at least one thing to be true: the maxim "Women and children first" held true in the case of the Titanic.